---
title: "Featured topic: International donor funding for tuberculosis: an analysis of donor reports to the Organisation for Economic Co-operation and Development"
author: "Peter Nguhiu, Takuya Yamanaka"
date: "`r Sys.Date()`"
knit: (function(inputFile, encoding) {
      out_dir <- "./local";
      rmarkdown::render(inputFile,
                        encoding=encoding,
                        output_dir=file.path(dirname(inputFile), out_dir))}) # added this to define a location for saving outputs
output: 
  html_fragment:
    # Do not include a table of contents
    toc: no
    # Set standard figure width to 12 inches
    fig_width: 12
    # Do not write figure captions
    fig_caption: FALSE
    # to save intermediate .png
    keep_md: true
    # Don't embed external resources (stylesheets, JS libraries) in the output 
    self_contained: False    
    # Don't include <div> tags around a header and the content found until the next header
    section_divs: FALSE  
---
```{r, echo=FALSE, warning=FALSE, message=FALSE}

# Set chunk options.
# Results "asis" is useful to output markdown from a function
# Suppress messages, warnings and also the ## at the beginning of printed text
knitr::opts_chunk$set(echo = FALSE, 
                      results = "asis",
                      message = FALSE,
                      warning = FALSE)
# Set output folder -- can only do this if not using the 
# RStudio knit button and instead doing the following from the 
# command line:
# rmarkdown::render("ch5_txt.rmd", output_file = "ch5.html")
# Establish the report year ----
# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
report_year <- 2023
# And the latest year for which OECD data are being displayed in graphics. Always two years older than report year
latest_year <- 2021
# Kill any attempt at using factors, unless we explicitly want them!
options(stringsAsFactors=FALSE)

# Load R functions ----
# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
source(here::here("report/functions/html_links.R"))
source(here::here("report/functions/output_ggplot.R"))

# Set up the running environment ----
# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
#source("io/set_environment.r")  # particular to each person so this file is in the ignore list
# Load packages ----
# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
library(ggplot2)
library(stringr)
library(dplyr)
library(tidyr)
library(RColorBrewer)
library(gridExtra)
library(gtbreport) 
library(jsonlite)

library(treemapify) 
library(pals)


outputs_folder_name = "report/ch4_data"

# Load donor data from OECD - summary objects only ----
# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
# Country info
source(here::here('report/ft_oecd_prepare_data.R'))
country_info <- haven::read_dta(here::here(base_folder,data_folder_name,"country_info.dta"))

# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

# Show static chart in addition to Kendo chart?
show_static = F

# Save underlying data files as CSV and charts as PDF files?
pdf_csv_folder = here::here("report/local/figures/ft_oecd") # Local folder only
save_csv = TRUE
save_pdf = TRUE

# Create the output folder (only if it doesn't yet exist)
dir.create(pdf_csv_folder, showWarnings = FALSE, recursive = TRUE)

```

```{r css_js}
# Add standard stylesheets and javascript to support kendo
cat(writeLines(readLines(here::here("report/resources/headers.htm"))))
```


# International donor funding for tuberculosis: an analysis of donor reports to the Organisation for Economic Co-operation and Development

<br>
In 2002, the World Health Organization (WHO) started annual monitoring of funding for tuberculosis (TB) prevention, diagnostic and treatment services, based on data reported by national TB programmes (NTPs) in annual rounds of global TB data collection. Findings have been published in global TB reports and peer-reviewed publications (`r ref_lnk("1&ndash;3")`). However, not all international donor funding for TB is captured in the data reported to WHO. Therefore, each year, WHO complements its analysis of data reported by NTPs with an assessment of international donor funding for TB, based on donor reports to the Organisation for Economic Co-operation and Development (OECD).

The OECD’s creditor reporting system (CRS) (`r ref_lnk("4")`) is the most comprehensive source of information about international donor funding. The CRS Aid Activity database makes it possible to analyse where aid goes, what purposes it serves and what policies it aims to support, on a comparable basis for all members of the OECD Development Assistance Committee (DAC). Data are for developing countries or areas eligible to receive official development assistance (ODA); they are collected for individual projects and programmes, with a focus on financial data (`r ref_lnk("4")`). 

As of `r report_year`, funding data (both commitments and disbursements) were provided by 37 multilateral donor organizations, 30 members of the OECD’s DAC (29 individual countries and the European Union) and a further 20 countries beyond the DAC that report to the OECD. Disbursement data include both direct transfers to countries and the provision of goods and services (e.g. in-kind transfers or technical assistance). 

Data on gross total official disbursements for TB (code 12263: Tuberculosis control) received by non-OECD countries during 2013–`r latest_year` were analysed (as of July `r report_year`, the latest year for which data have been reported is `r latest_year`). The CRS does not capture funding for TB that flows from one OECD member to an institution or government within the OECD. In addition, government contributions that are channelled through multilateral organizations - such as the Global Fund to Fight AIDS, Tuberculosis and Malaria (the Global Fund) - are attributed to the multilateral organization and not to the government of origin.

`r lnk("Fig. 1")` shows trends in international donor funding between 2013 and `r latest_year` for the six WHO regions and for all low- and middle-income countries (LMICs). The total for LMICs from all sources in `r latest_year` was US\$&nbsp;`r all_donation` million.



### `r anch("Fig. 1")` <span class="fig" style="color:#F21905">Fig. 1</span> International donor funding for TB  by source based on donor reports to the OECD, 2013&ndash;`r latest_year`



``` {r fig1, fig.alt="International donor funding for TB prevention, diagnosis and treatment by source, globally and by WHO region", fig.height = 6}

 # summarise_at(.vars = "value", sum, na.rm = TRUE) %>% 
Fig1_plot <- Fig1 %>%
 ggplot(aes(x=year, y = value, col = donor)) +
 geom_line(size=1, alpha=.85) +
 facet_wrap(~g_whoregion, scales = 'free_y', strip.position = "top" , 
            labeller = label_wrap_gen(width = 25))+
 scale_y_continuous(name = paste0("Millions (constant ", report_year-2," US$)")) +
 scale_x_continuous("", breaks=seq(2013,latest_year,2)) +
 scale_color_manual(values = c("#000000","#4ABAFC","#E63E13",#"#37F546",
                               "#BCBCBC"),
                    show)+
 theme_gtb()

output_ggplot(Fig1_plot, Fig1_plot$data, show_static, pdf_csv_folder, save_csv, save_pdf)
```
<div id="fig_1_lmc"></div>

<div class="row">
<div class="col-md-4">
<div id="fig_1_afr"></div>
</div>
<div class="col-md-4">
<div id="fig_1_amr"></div>
</div>
<div class="col-md-4">
<div id="fig_1_sear"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_1_eur"></div>
</div>
<div class="col-md-4">
<div id="fig_1_emr"></div>
</div>
<div class="col-md-4">
<div id="fig_1_wpr"></div>
</div>
</div>

<div class="footnote">Global Fund: Global Fund to Fight AIDS, Tuberculosis and Malaria; OECD: Organisation for Economic Co-operation and Development; WHO: World Health Organization.<br>^a^ As per the 2023 World Bank classification. The subtotals for the six WHO regions do not sum to the total for all low and middle-income countries combined. This is because funding recorded in regional (rather than country-specific) categories used by the OECD cannot be matched to a specific WHO region, and the recipient country for country-level funding is sometimes not specified.</div>


<hr>
<br />
From 2013 to `r latest_year`, the Global Fund was consistently the largest provider of international donor funding in LMICs, with its share averaging 65% in this period. In `r latest_year`, the total was US\$&nbsp;`r gf_donation` million (`r round(100*gf_donation/all_donation,0)`% of the global total of US$&nbsp;`r all_donation` million). 

Annual disbursements of bilateral funding from the United States of America (US) between 2013 and 2021 that were recorded in the OECD database (`r lnk("Fig. 1")`) ranged from US\$&nbsp;`r ftb(Fig1 %>% filter(donor == "United States of America" & g_whoregion =="Low- and middle-income\ncountriesᵃ" & year==2016) %>% select(value) %>% unlist())` million to US\$&nbsp;`r ftb(Fig1 %>% filter(donor == "United States of America" & g_whoregion =="Low- and middle-income\ncountriesᵃ" & year==2014) %>% select(value) %>% unlist())` million; in 2021, the total amount disbursed was US\$&nbsp;`r ftb(Fig1 %>% filter(donor == "United States of America" & g_whoregion =="Low- and middle-income\ncountriesᵃ" & year==2021) %>% select(value) %>% unlist())` million. These amounts are lower than official US congressional allocations for TB (`r ref_lnk("5, 6")`), which have grown from US\$&nbsp;`r ftb(Fig2 %>% filter(year == 2013) %>% select(value) %>% unlist())` million in 2013 to US\$&nbsp;`r ftb(Fig2 %>% filter(year == 2021) %>% select(value) %>% unlist())` million in 2021, US\$&nbsp;`r ftb(Fig2 %>% filter(year == 2022) %>% select(value) %>% unlist())` million in 2022 and US\$&nbsp;`r ftb(Fig2 %>% filter(year == 2023) %>% select(value) %>% unlist())` million in 2023 (`r lnk("Fig. 2")`).   


### `r anch("Fig. 2")` <span class="fig" style="color:#F21905">Fig. 2</span> Official US congressional allocations of funding for TB, 2013&ndash;`r report_year`

``` {r fig2, fig.alt="US funding for TB"}

Fig2_plot <- Fig2   %>% 
 filter(year >= 2013 ) %>% 
 ggplot(aes(x=year, y = value)) +
 geom_bar(stat = "identity", fill = "#4ABAFC") +
 scale_y_continuous(name = paste0("Millions (current US$)")) +
 scale_x_continuous("", breaks=seq(2013,report_year,2)) +
 theme_gtb() +
 annotate("segment", x=-Inf, 
          xend=Inf, 
          y=-Inf, 
          yend=-Inf, 
          colour = "#BCBCBC") 

output_ggplot(Fig2_plot, Fig2_plot$data, show_static, pdf_csv_folder, save_csv, save_pdf)
```
<div id="fig_2"></div>

<hr>
<br />

When bilateral funding from individual countries reported to the OECD is combined with estimates of individual country contributions for TB that are channelled through the Global Fund, the US government is by far the largest source of international donor funding for TB (`r lnk("Fig. 3")`). In the period 2013&ndash;2021, it provided `r ftb(Fig3 %>% filter(donor == "United States of America") %>% select(pcnt) %>% unlist())`% of the total amount of donor funding for TB from entities that report to the OECD; in the most recent year for which data are available (2021), the share was `r ftb(us_trend %>% filter(year == latest_year) %>% select(us_pcnt) %>% unlist())`%. The next largest individual country contributors in the period 2013&ndash;2021 were (in order of their contribution) the United Kingdom of Great Britain and Northern Ireland (`r ftb(Fig3 %>% filter(donor == "United Kingdom") %>% select(pcnt) %>% unlist())`%), France (`r ftb(Fig3 %>% filter(donor == "France") %>% select(pcnt) %>% unlist())`%), Germany (`r ftb(Fig3 %>% filter(donor == "Germany") %>% select(pcnt) %>% unlist())`%), Japan (`r ftb(Fig3 %>% filter(donor == "Japan") %>% select(pcnt) %>% unlist())`%) and Canada (`r ftb(Fig3 %>% filter(donor == "Canada") %>% select(pcnt) %>% unlist())`%).  



### `r anch("Fig. 3")` <span class="fig" style="color:#F21905">Fig. 3</span> Distribution of international donor funding for TB when Global Fund contributions are allocated to individual countries and other reporting entities,^a^ 2013&ndash;`r latest_year`

<div class="subhead">Funding totals for each entity are the sum of a) funding reported to the OECD, excluding that channelled via the Global Fund and b) an estimate of the funding provided for TB via the Global Fund.^b^</div> 

``` {r fig3 , fig.alt="International donor funding (in US$ millions) for TB prevention, diagnosis and treatment from individual countries", fig.height = 8, , fig.width = 12}

Fig3_plot <- ggplot2::ggplot(Fig3,aes(area=pcnt,fill=donor,label=donor))+ 
  treemapify::geom_treemap(layout="squarified", col="white")+ 
  geom_treemap_text(place = "centre",layout = "squarified", size = 15,
                    fontface = "bold",
                    colour = Fig3$text_col) + 
  theme_gtb() + 
  theme(legend.position = "none") +
  scale_fill_manual(values  = Fig3$fill_col)
  # scale_fill_manual(values  = as.vector(rev(pals::brewer.ylgn(11))))

output_ggplot(Fig3_plot, Fig3, show_static = T, pdf_csv_folder, save_csv, save_pdf)


```
<div class="footnote">Global Fund: Global Fund to Fight AIDS, Tuberculosis and Malaria; United Kingdom: United Kingdom of Great Britain and Northern Ireland; WHO: World Health Organization.<br>
^a^ Other reporting entities includes any countries beyond the 10 that are labelled that provide international donor funding for TB, and multilateral agencies (excluding the Global Fund).<br>
^b^ The estimate is based on the assumption that the share of a country or other entity’s contribution to TB funding provided by the Global Fund is the same as its share of total contributions to the Global Fund. For example, if a country provided 5% of the total contributions to the Global Fund, it was assumed to provide 5% of the TB funding attributed to the Global Fund in the OECD database.</div>


<hr>
<br />

`r lnk("Fig. 4")` shows that international funding for TB (US\$&nbsp;`r 1000 * (Fig4 %>% filter(year == latest_year & programme == "TB control") %>% select(value) %>% unlist() %>% round(3))` million in `r latest_year`) is about a third of that for malaria (US\$&nbsp;`r (Fig4 %>% filter(year == latest_year & programme == "Malaria control") %>% select(value) %>% unlist() %>% round(1))` billion in `r latest_year`) and about 13% of that for HIV (US\$&nbsp;`r (Fig4 %>% filter(year == latest_year & programme == "STD control including HIV") %>% select(value) %>% unlist() %>% round(1))` billion in `r latest_year`). Overall, TB funding constitutes 2.7% of total ODA funding for health and population programmes (US$&nbsp;`r (Fig4 %>% filter(year == latest_year & programme == "Total health and population programmes") %>% select(value) %>% unlist() %>% round(0))` billion in `r latest_year`).



### `r anch("Fig. 4")` <span class="fig" style="color:#F21905">Fig. 4</span> International donor funding for Health and Population programmes, TB, HIV and malaria, 2013&ndash;`r latest_year`
 

``` {r fig4 , fig.alt="Funding for TB compared with global target"}

# Create a label dataset (since the graph is faceted and easiest way to label is to create this)
# graph_labels <- FigB5.3 %>% select(x=year, y = value, name) %>% 
#  filter(x %in% c(2010, 2013, 2017, latest_year))


Fig4_plot <- Fig4 %>% 
 ggplot(aes(x=year, y = value, col = programme)) +
 geom_line(size=1, alpha=.85) +
 # geom_text(data = graph_labels, aes(x,y=y+1,label=signif(y,2)), 
 #           col = "#222222",vjust = "top", hjust = "middle") +
 scale_y_continuous(name = paste0("Billions (constant ", report_year-2," US$)")) +
 scale_x_continuous("", breaks=seq(2013,latest_year,2)) +
 scale_color_manual(values = c("#4ABAFC","#E63E13","#37F546", "#000000"),
                    show)+
 theme_gtb(
  #base_size = 16, axis_text_size = 12, axis_text_face = "bold"
 )+
 # theme(legend.position = "none") +
 # Add line labels
 # annotate("text", x = latest_year-1, y = 0, label = "TB", col="#4ABAFC") +
 # annotate("text", x = latest_year-1, y = 5, label = "HIV", col="#E63E13") +
 # annotate("text", x = latest_year-1, y = 3, label = "Malaria", col="#37F546") +
 # annotate("text", x = latest_year-1, y = 21, label = "Health", col="#000000") +
 # 
 annotate("segment", x=-Inf, 
          xend=Inf, 
          y=-Inf, 
          yend=-Inf, 
          colour = "#BCBCBC") 

output_ggplot(Fig4_plot, Fig4, show_static, pdf_csv_folder, save_csv, save_pdf)
```
<div id="fig_4"></div>
<div class = "footnote">HIV: human immunodeficiency virus; STD: sexually transmitted disease. </div>
<br />



`r anch("refs")`

<hr style="border:1px solid gray20">

**References**

1.	Floyd K, Fitzpatrick C, Pantoja A, Raviglione M. Domestic and donor financing for tuberculosis care and control in low-income and middle-income countries: an analysis of trends, 2002–11, and requirements to meet 2015 targets. Lancet Glob Health. 2013;1(2):e105–15. doi: https://doi.org/10.1016/S2214-109X(13)70032-9.

2.	Floyd K, Pantoja A, Dye C. Financing tuberculosis control: the role of a global financial monitoring system. Bull World Health Organ. 2007;85(5):334–40. doi: https://doi.org/10.2471/blt.06.034942.

3.	Su Y, Baena IG, Harle AC, Crosby SW, Micah AE, Siroka A et al. Tracking total spending on tuberculosis by source and function in 135 low-income and middle-income countries, 2000–17: a financial modelling study. Lancet Infect Dis. 2020;20(8):929–42. doi: https://doi.org/10.1016/S1473-3099(20)30124-9.

4.	OECD.Stat [website]. Paris: Organisation for Economic Co-operation and Development; 2022 (https://stats.oecd.org).

5.	The US government and global tuberculosis efforts. San Francisco: Kaiser Family Foundation; 2023 (https://www.kff.org/global-health-policy/fact-sheet/the-u-s-government-and-global-tuberculosis-efforts/?msclkid=2a614d82caf311ecb251baaa5c42217b).

6.	US foreign assistance dashboard [website]. Washington, DC: Office of Management and Budget, Agency Congressional Budget Justifications, Congressional Appropriation Bills; 2023 (https://www.foreignassistance.gov/).


<script type="text/javascript">
/* JSON data objects for the figures */

var fig_OECD1_data = `r Fig1 %>% pivot_wider(names_from = donor, values_from = value) %>% rename(tot=3,gf=4,us=5,oth=6) %>% toJSON("rows") `; 

var fig_OECD2_data = `r Fig2 %>% toJSON("rows") `; 

var fig_OECD4_data = `r Fig4 %>% pivot_wider(names_from = programme, values_from = value) %>% rename(malaria=3,tb=4,std=5,total=6) %>% toJSON("rows") `; 

</script>
```{r js_functions}
# Insert javascript file containing common Kendo number formatting functions ----
cat(writeLines(readLines(here::here("report/resources/gtbr_js.htm"))))
```

```{js, echo=FALSE}

/* Functions to create the figures */

function createFig_OECD2(fig_ID, data) {

  	// Filter the dataset on the country variable
		$(fig_ID).kendoChart({
			dataSource: data,
			chartArea: {
				height: 300
			},	

			legend: {
				position: "bottom",
				orientation: "vertical",
			},
			seriesDefaults: {
				type: "column"
			},
			series: [
			{name: "",
				field: "value",
				color: "dodgerblue",
        tooltip: {
				visible: true,
				template: "#= category #: US$ #= num_spacer(value) # million"
			}
			},],
			valueAxis: {
				title: {
					text: "Millions (current US$)"
				},
				line: {
					visible: false
				},
			},
			categoryAxis: {
				field: "year",
				labels: {
					rotation: 0,
          step: 2
				},
				majorGridLines: {
					visible: false
				},
				title: {
					text: ""
				}
			},
		});
}



function createFig_OECD1(fig_ID,data,filter) {
  	// Filter the dataset on the country variable
		dataJSON = data.filter( element => element.g_whoregion == filter);
  
		$(fig_ID).kendoChart({		
      dataSource: dataJSON,			
			chartArea: {
				height: 400
			},	     
      title: {
				text: filter,
				color: "black",
				font: "bold 14px  Arial,Helvetica,sans-serif",
        align: "center"
			},	  
			legend: {
				position: "bottom"
			},
			seriesDefaults: {
				type: "line",
			},
			series: [{
        name: "Total",
				field: "tot",
				color: "#000000",
        markers: {
          size: 5
        },
        tooltip: {
				visible: true,
				template: "Development assistance, total (#= category #): #= num_spacer(value) # million (constant 2021 US$)"
			}
			},{
        name: "Global Fund",
				field: "gf",
				color: "#4ABAFC",
        markers: {
          size: 5
        },
        tooltip: {
				visible: true,
				template: "Global Fund (#= category #): #= num_spacer(value) # million (constant 2021 US$)"
			}
			},{
        name: "United States of America",
				field: "us",
				color: "#E63E13",
        markers: {
          size: 5
        },
        tooltip: {
				visible: true,
				template: "United States of America (#= category #): #= num_spacer(value) # million (constant 2021 US$)"
			}
			},{
        name: "Other",
				field: "oth",
				color: "#BCBCBC",
        markers: {
          size: 5
        },
        tooltip: {
				visible: true,
				template: "Other (#= category #): #= num_spacer(value) # million (constant 2021 US$)"
			}
			},],
			valueAxis: {
				labels: {
				template: "#= kendo.format('{0}',value) #",
				},
				title: {
					text: "Millions (constant 2021 US$)"
				},
				line: {
					visible: false
				},
			},
			categoryAxis: {
				field: "year",
				labels: {
					rotation: 0,
					step: 2
				},
				majorGridLines: {
					visible: false
				}			}
		});
}


function createFig_OECD4(fig_ID,data) {
  	// Filter the dataset on the country variable

		$(fig_ID).kendoChart({		
      dataSource: data,			
			chartArea: {
				height: 600
			},	     
			legend: {
				position: "bottom"
			},
			seriesDefaults: {
				type: "line",
			},
			series: [{
        name: "TB control",
				field: "tb",
				color: "#4ABAFC",
        markers: {
          size: 5
        },
        tooltip: {
				visible: true,
				template: "TB control (#= category #): #= value.toPrecision(2) # billion (constant 2021 US$)"
			}
			},{
        name: "STD control including HIV",
				field: "std",
				color: "#E63E13",
        markers: {
          size: 5
        },
        tooltip: {
				visible: true,
				template: "STD control including HIV (#= category #): #= value.toPrecision(2) # billion (constant 2021 US$)"
			}
			},{
        name: "Malaria control",
				field: "malaria",
				color: "#37F546",
        markers: {
          size: 5
        },
        tooltip: {
				visible: true,
				template: "Malaria control (#= category #): #= value.toPrecision(2) # billion (constant 2021 US$)"
			}
			},{
        name: "Total health and population programmes",
				field: "total",
				color: "#000000",
        markers: {
          size: 5
        },
        tooltip: {
				visible: true,
				template: "Total health and population programmes (#= category #): #= value.toPrecision(2) # billion (constant 2021 US$)"
			}
			},],
			valueAxis: {
				labels: {
				template: "#= kendo.format('{0}',value) #",
				},
				title: {
					text: "Billions (constant 2021 US$)"
				},
				line: {
					visible: false
				},
			},
			categoryAxis: {
				field: "year",
				labels: {
					rotation: 0,
					step: 2
				},
				majorGridLines: {
					visible: false
				}			}
		});
}


```

```{js, echo=FALSE}

/* Create the figures after the document has been loaded */
$(document).ready(function() {
  createFig_OECD1("#fig_1_lmc",fig_OECD1_data,"Low- and middle-income\ncountriesᵃ");
  createFig_OECD1("#fig_1_afr",fig_OECD1_data,"WHO African Region");
  createFig_OECD1("#fig_1_amr",fig_OECD1_data,"WHO Region of\nthe Americas");
  createFig_OECD1("#fig_1_sear",fig_OECD1_data,"WHO South-East Asia\nRegion");
  createFig_OECD1("#fig_1_eur",fig_OECD1_data,"WHO European Region");
  createFig_OECD1("#fig_1_emr",fig_OECD1_data,"WHO Eastern Mediterranean\nRegion");
  createFig_OECD1("#fig_1_wpr",fig_OECD1_data,"WHO Western Pacific\nRegion");
  createFig_OECD2("#fig_2",fig_OECD2_data);
  createFig_OECD4("#fig_4",fig_OECD4_data);
  
});

```
